---
output: rmarkdown::github_document
bibliography: README-refs.bib
---

# Deep Time-Series Learning

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

> Estimation and Inference for Causal Effects with Single Binary Time Series

This repo provides basic implementation of the targeted maximum likelihood estimation of causal
effects based on the observation of a single binary time series. The approach considers the case where at each time point in a time series, a covariate vector, treatment, and outcome are observed in
chronological order. A family of causal effects are defined as the impacts of
stochastic interventions on a subset of the treatment nodes on a future outcome.
This general formulation of the statistical estimation problem subsumes many
other important estimation problems, including but not limited to classical time
series models, group sequential adaptive designs, and even independent and
identically distributed data where the summary measure of the past is simply the
empty set. For a general introduction to the targeted learning methodology and statistical causal
inference, please consult @vdl2011targeted and @vdl2017targeted. For details on the theoretical underpinnings of the approach, the interested reader may consider consulting @vdl2017online to start. For the actual R package implementing a more general methodology and other time-series based target parameters, see [tstmle](https://github.com/podTockom/tstmle/).  

---

## Example

To illustrate how to ascertain the effect of an
intervention on a single binary time series, consider the following example:

```{r example}
# Forthcoming...check back later.
```

---

## References

