---
title: "TMLE of Causal Effects of Interventions on a Single Binary Time Series"
author: "Ivana Malenica"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: intro_tstmle01.bib
vignette: >
  %\VignetteIndexEntry{TMLE of Causal Effects of Interventions on a Single Binary Time Series}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE}
options(warn = FALSE)
```

## Introduction

The `tstmle01` package implements targeted maximum likelihood estimation (TMLE) of the marginal causal effect based on observation of a single binary time-series [@c1]. Current implementation supports iterative TMLE; check back for one-step and online TMLE [@c4], [@c5]. 

For the (faster) package implementing a more general methodology and other time-series based target parameters, see [tstmle](https://github.com/podTockom/tstmle/). In particular, [tstmle](https://github.com/podTockom/tstmle/) implements the data-dependent, $C_{o(t)}-$specific, causal effect and the adaptive design for learning the optimal treatment rule within a single time series [@c2]. Here, initial estimation is based on the [sl3](https://github.com/jeremyrcoyle/sl3) package, which constructs ensemble models with proven optimality properties for time-series data [@c3].

We emphasize that this general formulation of the statistical estimation problem subsumes many
other important estimation problems, including but not limited to classical time
series models, group sequential adaptive designs, and even independent and
identically distributed data when the summary measure of the past is simply the
empty set. 

---

## General methodology

We consider the case where we observe a single sequence of dependent random
variables $O(1), \dots O(n)$, where each $O(t)$ with $t \in \{1, \dots n\}$ takes values in $\mathbf{R}^p$.
Further, we assume that at each time $t$, we have a chronological order of the
covariate vector $W(t)$, treatment or exposure $A(t)$ and outcome of interest
$Y(t)$. We impose the conditional (strong) stationarity assumption on the
dependent process in question and limit the statistical model to
conditionally stationary distributions in which widely-separated observations
are asymptotically independent. We propose a causal model that is compatible
with this statistical model, and define a family of causal effects in terms of
stochastic interventions on a subset of the treatment nodes. We note that static interventions are just a subset of the more general stochastic interventions. Our statistical target parameter is defined as the mapping $\Psi_{g^*}^n : \mathcal{M}^n \rightarrow \mathbb{R}$ such that:

$$
\Psi_{g^*}^n(P^n) = \Psi_{g^*}^n(p_W, p_A, p_Y) = \mathbb{E}_{P_{g^*}^n}
 Y_{g^*}(\tau_{g^*})
$$

where $Y_{g^*}(\tau_{g^*})$ is the outcome at time $\tau$ for the intervened data $O_{g^*}^n$. Therefore, we are interested in the expectation of $Y(\tau_{g^*})$ under the intervened distribution $P_{g^*}^n$. 

We can also define a causal effect relative to some baseline level of treatment as
follows:

$$
\Psi^n(P^n) = \Psi_{g^*}^n(P^n) - \Psi_{g_0^*}^n(P^n),
$$

where $g_0^*$ is a baseline treatment regime on the intervention nodes $A(\mathcal{I}) = (A(t) : t \in \mathcal{I})$.

A key feature of the estimation problem is that data is dependent and statistical inference is based on asymptotics in time.This is a unique and powerful result as it estimates the causal effect based on a single time series, and without the need for a control time series. 

---

## Installing and loading the package 

In the following sections, we examine the use of `tstmle01` in a variety of simple examples. The package can be installed as follows:

```{r install_pkg}
if (!("tstmle01" %in% installed.packages())) {
  devtools::install_github("podTockom/tstmle01")
}
```

Once the package is installed, we can load it using the following command:

```{r load_pkg}
suppressMessages(library(tstmle01))
```

---

## Input Data

Input data should be a single time-series of arbitrary size. The longer your provided time-series, better the estimation process, just like in the case of $n$ i.i.d. samples (alas, longer computational demand). The input should be $n$ by $1$ \code{data.frame} with row names indicating which node that particular time point belongs to: $W$, $A$, or $Y$. As mentioned in the previous section, $W$ are the set of covariates, $A$ intervention nodes, and $Y$ outcome. The imposed order must be $W$, $A$, $Y$, so that each time point will be a set of $(W,A,Y)$ for $i = 1, ..., n$. For more details, see the example simulated dataset in the data directory. 

```{r dataFormat}
data("ts_samp_data")
head(ts_samp_data)
```

## Single intervention 

In this section we examine the effect of a single time-point intervention on a single outcome upstream.
In particular, we intervene on time point $A(3)$ by setting it equal to $1$ with probability $1$, and examine the expected value of $Y(5)$ under this intervened distribution. This simple example is also a special case of a static intervention on node $A(3)$; we could have specified any probability of success here. In order to save time, for all examples considered here we use a shortened time-series located in the data directory. For estimation purposes, we use 100 Monte Carlo draws to estimate conditional means of $Y_{g^*}(\tau_{g^*})$ and 100 Monte Carlo draws to estimate both $h^*$ and $h$-densities. 

Note that there is a direct relationship between the size of a time-series and the outcome time point one might be interested in. In particular, for the clever covariate to be properly calculated, one must be aware of how long of a time-series we want to use for estimation purposes and where our outcome value is in the time-series. Since we are calculating expected value of $Y(\tau)$ with different lags from $i=1$ to $i=N$, if we want $\tau = 5$, we will need our input to be at least $N+4$. Having that in mind, later outcomes of interest will need longer time-series and will take longer to compute. Therefore, if one is concerned with the effect of an intervention at a specific $\tau$, we advise to excise parts of the time-series with $\tau$ being towards the beginning of the time-series.

```{r singleInter, eval=FALSE}
#Load package:
suppressMessages(library(tstmle01))

#Load sample data:
data(ts_samp_data_ex)
data<-ts_samp_data_ex

# Get initial estimates of the process:
fit <- initEst(data, t=5, freqW = 2, freqA = 2, freqY = 2)

#Get targeted estimate of Y(\tau) under intervention P(A(3)=1|C_A(3))=1:
est1 <- mainTMLE(fit, t = 5, Anode = 3, intervention = 1, B = 100, N = 100, MC = 100)

```


## Session Information

```{r sessionInfo, echo=FALSE}
sessionInfo()
```

---

## References

